---
title: "SWAT Raab - SoilGrids setup for clustering"
output:
  html_notebook: default
  html_document: default
---

###Input data

Following input data needs to be place in the working directory of this script

File/Folder   									Location
-----         									-----
Uncompressed "clValid" folder   				/root
"SoilGrid250m" folder with TIFF files   		/root
Empty folder "SoilGrid250m" for output files	/root/01_Datengrundlage/04_SOIL/
Shape file subs1.shp							/root


###Libraries
To perform the unsupervised machine learning clustering the "optCluster" package needs to be installed. This package requires another one called "clValid", which was modified to improve the clustering of SOM, from unidimensional maps (1 x number of clusters) to rectangular or quadratic ones. Therefore this package "clValid" needs to be install before the installation of "optCluster".

```{r}
# install.packages("clValid", repos = NULL, type="source")

require(euptf)
require(raster)
require(rgdal)
require(sp)
require(stats)
require(optCluster)
library(pasta)
library(magrittr)
library(dplyr)
library(tibble)
```

###Preparation of spatial input file

```{r}

rst_dir <- "E:/Mostviertel/soilgrids"
wrt_dir <- "E:/Mostviertel/sg_kmeans"

#-------------------------------------------------------------------------#
#Loading the Basin boundary as extent for creation of raster subsets from
# the tiles.
#-------------------------------------------------------------------------#
shp_file <- "E:/Mostviertel/DEM/boundary.shp"
bnd_shp <- readOGR(shp_file, layer = "boundary")
#-------------------------------------------------------------------------#
#Loading all GEO TIFF for the tile covering Austria.
#-------------------------------------------------------------------------#
tile_list <-  list.files(rst_dir, pattern = "_250m.tif$")

rst_list <- list()

select_label <- function(text) {
  text[text %in% c("BDRICM", "BLDFIE", "CECSOL", "CLYPPT", 
                   "CRFVOL", "ORCDRC", "PHIHOX", "SLTPPT", 
                   "SNDPPT", "sl"%&%1:7)]
}

set_nodata <- function(rst) {
  if (rst@data@min == 0 & rst@data@max == 255) {
    rst@file@nodatavalue <- 255
  } else {
    rst@file@nodatavalue <- -32768
  }
  return(rst)
}

for(i_rst in tile_list) { 
  i_name <- i_rst %>% 
    strsplit(., "_") %>% 
    unlist() %>% 
    select_label(.) %>% 
    substr(., 1, 3) %>% 
    paste(., collapse = "_")
  rst_tmp <- raster(rst_dir%//%i_rst) %>% 
    set_nodata(.) %>% 
    projectRaster(., crs = crs(bnd_shp)) %>% 
    crop(., extent(bnd_shp)) %>% 
    mask(., bnd_shp) %>% 
    as.data.frame() %>% 
    filter(!is.na(.[,1])) %>% 
    set_colnames(i_name)
  rst_list[[i_name]] <- rst_tmp
}

rst_tbl <- rst_list %>% 
  bind_cols() %>% 
  as_tibble() %>% 
  set_colnames(tolower(colnames(.)))

clst_tbl <- rst_tbl %>% 
  dplyr::select(ends_with("sl1"), 
         ends_with("sl3"), 
         ends_with("sl6"), 
         ends_with("bdr")) %>% 
  mutate_at(vars(starts_with("orc")), funs(./10)) %>% 
  mutate_at(vars(starts_with("bld")), funs(./1000)) %>% 
  mutate_at(vars(starts_with("phi")), funs(./10)) %>% 
  scale(., scale = TRUE, center = TRUE) %>% 
  as.data.frame() %>% 
  set_rownames("cell"%_%1:nrow(.))

save(rst_tbl, file = wrt_dir%//%"rst_tbl.RData")
load(wrt_dir%//%"rst_tbl.RData")

rst_km <- list()
wss <- (nrow(clst_tbl)-1)*sum(apply(clst_tbl,2,var))

for(i in 1:20) {
  rst_km[["n"%_%i]] <- kmeans(x = clst_tbl,centers = i, iter.max = 100)
  wss[i] <- sum(rst_km[["n"%_%i]]$withinss)
}

plot(1:20, wss, type="b", xlab="Number of Clusters", ylab="Within groups sum of squares")

```

```{r}
#Best choice here (when very low numbers are excluded) is 8 classes.

clust_14 <- rst_km$n_2$cluster %>% 
  enframe() %>% 
  mutate(name = name %>% gsub("cell_", "", .) %>% as.numeric(.))

# Selected 14 classes


rst_dir <- "E:/Mostviertel/soilgrids"

bnd_file <- "E:/Mostviertel/DEM/boundary.shp"

bnd_shp <- readOGR(bnd_file, layer = "boundary")

rst_i <- raster(rst_dir%//%"BLDFIE_M_sl1_250m.tif") %>% 
  set_nodata(.) %>% 
  projectRaster(crs = crs(bnd_shp))

rst_i_c <- crop(rst_i, extent(bnd_shp)) %>% 
  mask(., bnd_shp)

len_rst <- length(rst_i_c)
dim_rst <- dim(rst_i_c)

clust_vct <- rep(NA, len_rst)
clust_vct[which(!is.na(rst_i_c@data@values))] <- clust_14$value

clust_rst <- clust_vct %>% 
  matrix(ncol = dim_rst[1], nrow = dim_rst[2]) %>% 
  t() %>% 
  raster(crs = crs(bnd_shp))

extent(clust_rst) <- extent(bnd_shp)

clust_g <- as(clust_rst, 'SpatialGridDataFrame')
writeGDAL(dataset = clust_g, fname =  wrt_dir%//%"sw_soil.tif", 
          drivername = "GTiff", type = "Int16")


# Load clip raster input files and and arrange in table
tif_pth <- rst_dir

tif_lst <- list.files(path = tif_pth, pattern = ".tif$")

rst_layer <- list()
for (i in tif_lst){
  name_i <- i %>% strsplit("_") %>% unlist()
  if(length(name_i) == 5){
    name_i <- substr(name_i[2], 1, 3)%_%substr(name_i[4],3,3)
    rst_layer[[name_i]] <- raster(tif_pth%//%i) %>% 
      projectRaster(crs = crs(bnd_shp)) %>% 
      crop(., extent(bnd_shp)) %>% 
      mask(., bnd_shp) %>% 
      as.data.frame() %>% 
      filter(!is.na(.[,1]))
  } else {
    name_i <- "zmax"
    rst_layer[[name_i]] <- raster(tif_pth%//%i) %>% 
      projectRaster(crs = crs(bnd_shp)) %>% 
      crop(., extent(bnd_shp)) %>% 
      mask(., bnd_shp) %>% 
      as.data.frame() %>% 
      filter(!is.na(.[,1]))
  }
  }

rst_tbl <- lapply(rst_layer, function(x){x})
for(i in names(rst_tbl)) rst_tbl[[i]] %<>% set_colnames(i)

sol_lyr <- list()

for(i_lyr in 1:7){
  sol_lyr[["lyr"%_%i_lyr]] <- rst_tbl %>% 
    bind_cols() %>% 
    dplyr::select(ends_with(as.character(i_lyr))) %>% 
    set_colnames(substr(colnames(.),1,3)) %>% 
    mutate(TOPSOIL = ifelse(i_lyr <= 4, "top", "sub")) %>% 
    as_tibble()
}

calc_solpar <- function(tbl) {
  tbl %>% 
    mutate(USSAND = snd,
           USSILT = slt,
           USCLAY  = cly,
           OC      = orc/10,
           BD      = bld/1000,
           CRF     = crf,
           PH_H2O  = phi/10,
           CEC     = cec) %>% 
    dplyr::select(TOPSOIL, USSAND, USSILT, USCLAY, OC, BD, CRF, PH_H2O, CEC) %>% 
    mutate(th_s  = predict.ptf(., ptf = "PTF06"),
           th_fc = predict.ptf(., ptf = "PTF09"),
           th_wp = predict.ptf(., ptf = "PTF12"),
           k_s   = predict.ptf(., ptf = "PTF17"),
           awc   = th_fc - th_wp) %>% 
    rename(snd = USSAND,
           slt = USSILT,
           cly = USCLAY,
           orc = OC,
           bld = BD,
           crf = CRF,
           phi = PH_H2O,
           cec = CEC) %>% 
    dplyr::select(-TOPSOIL)
}

sol_lyr %<>% lapply(., calc_solpar)

sol_lyr$lyr_0_30   <- ((sol_lyr$lyr_1*2.5 + sol_lyr$lyr_2*7.5 + 
                        sol_lyr$lyr_3*12.5 + sol_lyr$lyr_4*7.5) / 30) %>% 
  as_tibble()

sol_lyr$lyr_30_100 <- ((sol_lyr$lyr_4*15 + sol_lyr$lyr_5*35 + 
                        sol_lyr$lyr_6*20) / 70) %>% 
  as_tibble()

sol_lyr$lyr_100_200 <- ((sol_lyr$lyr_6*50 + sol_lyr$lyr_7*50) / 100) %>% 
  as_tibble()
  

sol_lyr <- sol_lyr[c("lyr_0_30", "lyr_30_100", "lyr_100_200")]

sol_lyr %<>% lapply(., function(x){
  x %>% 
    add_column(class = clust_14$value) %>% 
    group_by(class) %>% 
    summarise_all(funs(mean)) %>% 
    mutate(tex = psd2classUS(snd, slt, cly, orc, option=TRUE))}) 

sol_lyr$lyr_0_30    %<>% add_column(z = 300)
sol_lyr$lyr_30_100  %<>% add_column(z = 1000)
sol_lyr$lyr_100_200 %<>% add_column(z = rst_tbl$zmax %>% 
                                          add_column(class = clust_14$value) %>% 
                                          filter(!is.na(class)) %>% 
                                          group_by(class) %>% 
                                          summarise_all(funs(mean)) %>% 
                                          .[[2]] %>% 
                                          multiply_by(10))

assign_hydgrp <- function(k_s){
  hyd_grp <- c("D","C","B","A")
  hyd_trs <- c(3.6,36,144,9999)
  lapply(k_s, function(x) hyd_grp[hyd_trs > x][1]) %>% unlist()
}

arrange_lyr_i <- function(lyr_tbl) {
  lyr_tbl %>% 
    dplyr::select(z, bld, awc, k_s, orc, cly, slt, snd, crf) %>% 
    mutate(alb    = 0.6/exp(0.4*orc),
           usle_k = ((0.2 + 0.3*exp(-0.256*snd * (1 - slt/100)))) * 
             ((slt/(cly + slt))^0.3) * 
             (1 - 0.0256*orc / (orc + exp(3.72 - 2.95*orc))) *
             ((1 - 0.7*(1 - snd/100) / 
             ((1 - snd/100) + exp(-5.51 + 22.9*(1 - snd/100))))),
           ec     = 0,
           k_s = (10^k_s)/2.4)
}

gen <- tibble(objectid  = 1:nrow(sol_lyr[[1]]),
              muid      = rep("", nrow(sol_lyr[[1]])),
              seqn      = rep("", nrow(sol_lyr[[1]])),
              snam      = "solclst"%&%sol_lyr[[1]]$class,
              s5id      = rep("", nrow(sol_lyr[[1]])),
              cmppct    = rep("", nrow(sol_lyr[[1]])),
              nlayers   = length(sol_lyr),
              hydgrp    = assign_hydgrp(sol_lyr$lyr_0_30$k_s),
              sol_zmx   = sol_lyr$lyr_100_200$z,
              anion_exc = 0.5, 
              sol_crk   = 0.5,
              texture   = sol_lyr[[1]]$tex%_%sol_lyr[[2]]$tex%_%
                          sol_lyr[[3]]$tex)

sol_lyr %<>% lapply(., arrange_lyr_i) %>% 
  lapply(., round, digits = 2)


sol_out <- bind_cols(gen, bind_cols(sol_lyr)) %>% 
  bind_cols(., data.frame(matrix(0, nrow = nrow(sol_lyr[[1]]),
                                    ncol = 152 - ncol(.))))

write.table(sol_out, file = wrt_dir%//%"usersoil.csv", quote = FALSE, 
          row.names = FALSE, col.names = FALSE, sep = ",")

sol_lkp <- tibble(VALUE = substr(gen$snam, 9, 11),
                  NAME  = gen$snam)

write.table(sol_lkp, file = wrt_dir%//%"sol_lkp.csv", quote = FALSE, 
            row.names = FALSE, col.names = TRUE, sep = ",")
```

